Manuscript Figures

### 2021-08-25 src/manuscript_figures.Rmd
drop Fig 3B - hazard ratio barplots with confidence intervals

  # compute hazard ratios for the top factor
  v <- vars[1] 
  dat <- merge(clin, M[,v,drop = F], by = 'row.names')
  rownames(dat) <- dat$Row.names
  dat$Row.names <- NULL
  dat <- get_surv_df(surv, dat, event, time)
  # compute hazard ratios 
  cox1 <- coxph(Surv(time,status) ~ ., 
          data=dat, x = TRUE)
  hr <- data.frame(summary(cox1)[['conf.int']]) # get hazard ratios with conf intervals
  hr <- hr[grep(v, rownames(hr)), c(1,3,4)]
  colnames(hr) <- c('hr', 'hr_low', 'hr_high')
  rownames(hr) <- v
  p1 <- survminer::ggforest(cox1, data = dat)

  # split variable into high/low to compute survival adjusted curves
  cp <- find_survival_cutpoint(dat[,v,drop=F], surv, event_col = event, duration_col = time)
  dat[,v] <- ifelse(dat[,v] > cp, paste(v, 'high'), paste(v, 'low'))
  cox2 <- coxph(Surv(time,status) ~ ., 
                data=dat, x = TRUE)
  p2 <- survminer::ggadjustedcurves(fit = cox2, data = dat, variable = v, method = 'average')

@@

  return(list('p1' = p1, 'p2' = p2, 'hr' = hr, 
              'Cindex_base' = rsf_stats$C_wo, 'Cindex_with_LFs' = rsf_stats$C_with))

  return(NULL)
})
parallel::stopCluster(cl)

plots <- sapply(simplify = F, names(surv.stats), function(x) {   
  p <- surv.stats[[x]][['p2']] 
  if(!is.null(p)) { 
    p + theme(axis.text.x = element_text(size = 8), 
              axis.text.y = element_text(size = 8), 
              axis.title.x = element_blank(), 
              axis.title.y = element_blank(), 
              legend.title = element_blank(),
              legend.position = c(0.5, 0.9),
              legend.text = element_text(size = 8), 
              legend.background = element_blank(), 
              legend.direction = 'vertical') 
    }
})
plots <- plots[!sapply(plots, is.null)]
p1 <- cowplot::plot_grid(plotlist = plots, nrow = 3, 
                   labels = sapply(strsplit(names(plots), '\\.'), function(x) x[1]), 
                   label_size = 8)
# plot hazard ratios for top LF per cancer type 
hr <- do.call(rbind, lapply(names(surv.stats), function(x) {
  if(is.null(surv.stats[[x]])) {return(NULL)}
  message(x)
  df <- surv.stats[[x]][['hr']]
  df$project <- unlist(strsplit(x, '\\.'))[[1]]
  return(df)
}))
hr$latent_factor <- rownames(hr)
# remove hr with Inf values
hr <- hr[!is.infinite(hr$hr_high),]

p2 <- ggplot(hr, aes(x = reorder(paste0(project, ' (', latent_factor, ')'), hr), y = log(hr))) +
  geom_bar(stat = 'identity') + 
  geom_errorbar(aes(ymin = log(hr_low), ymax = log(hr_high)), width = 0.2) + 
  labs(x = '', y = 'log(hazard-ratio) of top latent factor per cancer\n(adjusted for tumor stage + age + gender)')+
  coord_flip()

>>>>

  # compute adjusted survival curve for thb top factor
  v <- vars[1] 
  dat <- merge(clin, M[,v,drop = F], by = 'row.names')
  rownames(dat) <- dat$Row.names
  dat$Row.names <- NULL
  dat <- get_surv_df(surv, dat, event, time)

  # split variable into high/low to compute survival adjusted curves
  cp <- find_survival_cutpoint(dat[,v,drop=F], surv, event_col = event, duration_col = time)
  dat[,v] <- ifelse(dat[,v] > cp, paste(v, 'high'), paste(v, 'low'))
  cox2 <- coxph(Surv(time,status) ~ ., 
                data=dat, x = TRUE)
  p <- survminer::ggadjustedcurves(fit = cox2, data = dat, variable = v, method = 'average')

@@

  return(list('p' = p, 
              'Cindex_base' = rsf_stats$C_wo, 
              'Cindex_with_LFs' = rsf_stats$C_with))
  return(NULL)
})
parallel::stopCluster(cl)

# combine adjusted survival curve plots into a single plot
plots <- sapply(simplify = F, names(surv.stats), function(x) {   
  p <- surv.stats[[x]][['p']] 
  if(!is.null(p)) { 
    p + theme(axis.text.x = element_text(size = 8), 
              axis.text.y = element_text(size = 8), 
              axis.title.x = element_blank(), 
              axis.title.y = element_blank(), 
              legend.title = element_blank(),
              legend.position = c(0.5, 0.9),
              legend.text = element_text(size = 8), 
              legend.background = element_blank(), 
              legend.direction = 'vertical') 
    }
})
plots <- plots[!sapply(plots, is.null)]
p1 <- cowplot::plot_grid(plotlist = plots, nrow = 3, 
                   labels = sapply(strsplit(names(plots), '\\.'), function(x) x[1]), 
                   label_size = 8)
